---
title: "Phase 2 Comparison Graphs"
output:
  html_document:
    theme: readable
    highlight: 
    toc: yes
    toc_depth: 2
    toc_float:
      smooth_scroll: yes
    code_download: true
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
## Markdown Options
knitr::opts_chunk$set(echo = FALSE, error = FALSE, warning = FALSE, message = FALSE)
```

```{r}
#load packages
library(tidyverse)
library(ggplot2)
#library(ggtext) #for italics
library(RColorBrewer)
library(survival) # core survival analysis function
library(survminer) # recommended for visualizing survival curves
library(ggfortify)
```

# Load Data  {.hidden}

```{r}
#read CSVs
path_file <- "C:/Users/allie/OneDrive/Desktop/R Projects/HeatwaveProject/"
Data_All <- read_csv(paste0(path_file, "data/data_analysis/Data_All.csv")) %>% 
  mutate(Phase = as.factor(Phase))

Kestrel_AvgTemp_Sum <- read_csv(paste0(path_file, "data/data_QAQC/Kestrel_AvgTemp_Sum.csv"))
Kestrel_AvgTemp_Sum_Total <- read_csv(paste0(path_file, "data/data_QAQC/Kestrel_AvgTemp_Sum_Total.csv"))
```

# Plant Size Graphs

## Height by Basal Diameter

```{r}
Data_All %>% 
  group_by(Phase, Species) %>% 
  ggplot(aes(x = BasalDia_mm,
             y = Height_mm,
             color = Phase)) +
  geom_point(alpha = 0.03) +
  facet_wrap(~CommonName) +
  #geom_point(aes(size = Data$Biomass_g)) +
  xlim(2,11) +
  ylim(0,500) +
  xlab("Basal Stem Diameter") +
  ylab("Height (mm)") +
  labs(title = "Species Size Distribution") +
  scale_color_discrete(guide = guide_legend(override.aes = list(alpha = 1, size = 2))) +
  theme_minimal()
```

## Height

```{r}
#size boxplot
Data_All %>% 
  group_by(Phase, Species) %>% 
  ggplot(aes(x = Species,
             y = Height_mm,
             color = Phase)) +
  geom_boxplot() +
  ylim(0,600) +
  xlab("Basal Stem Diameter") +
  ylab("Height (mm)") +
  labs(title = "Species Size Distribution") +
  scale_color_discrete(guide = guide_legend(override.aes = list(alpha = 1, size = 2))) +
  theme_minimal()
```

## Basal Diameter

```{r}
#size boxplot
Data_All %>% 
  group_by(Phase, Species) %>% 
  ggplot(aes(x = Species,
             y = BasalDia_mm,
             color = Phase)) +
  geom_boxplot() +
  #ylim(0,.03) +
  xlab("Species") +
  ylab("Basal Diameter (mm)") +
  labs(title = "Species Size Distribution") +
  scale_color_discrete(guide = guide_legend(override.aes = list(alpha = 1, size = 2))) +
  theme_minimal()
```

## Age

```{r}
#size boxplot
Data_All %>% 
  group_by(Phase, Species) %>% 
  ggplot(aes(x = Species,
             y = Whorls,
             color = Phase)) +
  geom_boxplot() +
  ylim(0,8) +
  xlab("Species") +
  ylab("Whorls") +
  labs(title = "Species Size Distribution") +
  scale_color_discrete(guide = guide_legend(override.aes = list(alpha = 1, size = 2))) +
  theme_minimal()
```

# Temperature Graphs

```{r}


#add placeholder date to make graph
Kestrel_AvgTemp_Sum <- Kestrel_AvgTemp_Sum %>% 
  mutate(Date = as.Date("2021-10-15"))
Kestrel_AvgTemp_Sum$DateTime <- paste(Kestrel_AvgTemp_Sum$Date, " ", Kestrel_AvgTemp_Sum$Time)
Kestrel_AvgTemp_Sum$DateTime <- strptime(Kestrel_AvgTemp_Sum$DateTime, format="%Y-%m-%d %H%M")

Kestrel_AvgTemp_Sum$Phase <- as.factor(Kestrel_AvgTemp_Sum$Phase)
Kestrel_AvgTemp_Sum$Chamber <- as.factor(Kestrel_AvgTemp_Sum$Chamber)
```

## Temperature Fluctuations

```{r}
#All chambers
Kestrel_AvgTemp_Sum_graph <- Kestrel_AvgTemp_Sum %>% 
  filter(Kestrel == "actual", Heatwave == "no")

Kestrel_AvgTemp_Sum_graph %>% 
  group_by(Phase, Chamber, Kestrel, Heatwave) %>%
  ggplot(aes(x = as.POSIXct(DateTime),
              y = Temperature_avg,
              color = Chamber)) +
  geom_point() +
  facet_wrap(~Phase) +
  ylim(0, 40) +
  scale_x_datetime(date_labels = "%H%M") +
  xlab("Time") +
  ylab("Temperature (Celsius)") +
  labs(fill = "", tag = "(a)") +
  #labs(caption = "\nFIGURE S1a | Average Daily Temperatures Fluctuations of Growth Chambers") +
  theme_minimal() +
  theme(text = element_text(family = "serif", size = 10),
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 10),
        strip.text.x = element_text(size = 10),
        plot.caption = element_text(hjust = 0,
                                    family = "serif",
                                    #face = "bold",
                                    size = 10))
```

## Mean Temperatures

```{r}
#filter for actual temps
Kestrel_AvgTemp_Sum_Total_graph <- Kestrel_AvgTemp_Sum_Total %>% 
  filter(Kestrel == "actual")

#Graph!

#Bar graph
Kestrel_AvgTemp_Sum_Total_graph %>% 
  group_by(Phase, Heatwave) %>%
  ggplot(aes(x = Chamber,
             y = Temperature_avg,
             fill = Heatwave)) +
  geom_col(position= "dodge") +
  facet_wrap(~Phase) +
  ylim(0, 40) +
  ylab("Temperature (Celsius)") +
  labs(fill = "", tag = "(b)") +
  #labs(caption = "\nFIGURE S1b | Average Temperatures of Growth Chambers") +
  geom_text(label = Kestrel_AvgTemp_Sum_Total_graph$Temperature_avg,
            vjust = 1.5, position = position_dodge(0.9), color = "white") +
  #scale_fill_brewer(palette = "Set2") +
  scale_fill_discrete(direction = -1,
                      labels = c("Ambient", "Heatwave")) +
  theme_minimal() +
  theme(text = element_text(family = "serif", size = 10),
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 10),
        strip.text.x = element_text(size = 10),
        plot.caption = element_text(hjust = 0,
                                    family = "serif",
                                    #face = "bold",
                                    size = 10))
```

# Kaplan Meier Survival Curve

```{r}
#define custom color scale
myColorsPaired <- c("#6A3D9A", "#CAB2D6", "#FF7F00", "#FDBF6F",  "#33A02C", "#B2DF8A", "#E31A1C", "#FB9A99", "#1F78B4", "#A6CEE3")
myColorsDark <- c("#6A3D9A", "#FF7F00", "#33A02C", "#E31A1C", "#1F78B4")
myColorsLight <- c("#CAB2D6", "#FDBF6F", "#B2DF8A", "#FB9A99", "#A6CEE3")

# names(myColorsPaired) <- levels(Data_All_w$Legend)
# names(myColorsDark) <- levels(Data_All_w$Legend)
# names(myColorsLight) <- levels(Data_All_w$Legend)
# custom_colors <- scale_colour_manual(values = myColorsLight)
# custom_colors_fill <- scale_fill_manual(values = myColorsDark)

Data_All <- Data_All %>%
  mutate(Legend = Heatwave_graph)
#Phase 1 - Ambient
Data_All_amb <- Data_All %>% 
  filter(Phase == 1,
         Treatment_temp == "Ambient")
#Phase 1 - Ambient Heatwave
Data_All_amb_hw <- Data_All %>% 
  filter(Phase == 1,
         Treatment_temp == "Ambient_HW")
#Phase 1 - Combined graphs
Data_All_Phase1 <- Data_All %>% 
  filter(Phase == 1)
#Phase 2 - Hotter
Data_All_hot <- Data_All %>% 
  filter(Phase == 2,
         Treatment_temp == "Hotter")
#Phase 2 - Hotter Heatwave
Data_All_hot_hw <- Data_All %>% 
  filter(Phase == 2,
         Treatment_temp == "Hotter_HW")
#Phase 2 - Combined graphs
Data_All_Phase2 <- Data_All %>% 
  filter(Phase == 2)
```

## Phase 1 - Ambient

### Ambient  {.hidden}

```{r}
# KM curve - Ambient

#define colors
#names(myColorsPaired) <- levels(Data_All_amb$Legend)
names(myColorsDark) <- levels(Data_All_amb$Legend)
names(myColorsLight) <- levels(Data_All_amb$Legend)
custom_colors <- scale_colour_manual(values = myColorsLight)
custom_colors_fill <- scale_fill_manual(values = myColorsDark)

#Kaplan Meier Survival Curve - combined
km_species_fit <- survfit(Surv(Week, Dead_Count)~Legend, data = Data_All_amb)

autoplot(km_species_fit) +
  scale_x_continuous(limits = c(0, 36), breaks = seq(0 , 36, by = 2)) +
  xlab("Weeks") +
  ylab("Survival Probability") +
  # geom_text(label = "ponderosa pine",
  #           x = 12, y = 0.05, color = "black", size = 3, family = "serif") +
  # geom_text(label = "limber pine",
  #           x = 33, y = 0.05, color = "black", size = 3, family = "serif") +
  # geom_text(label = "Pinus ponderosa",
  #           x = 12, y = 0.05, color = "black", size = 3, family = "serif", fontface = "italic") +
  # geom_text(label = "Pinus flexilis",
  #           x = 33, y = 0.05, color = "black", size = 3, family = "serif", fontface = "italic") +
  labs(color = "", fill = "", tag = "(a)") +
  #labs(caption = "\nFIGURE 1 | Kaplan Meier Survival Curve of Juvenile Survival Probability under Drought.\nCurves shows the survival probability of each species under droughted and ambient temperature\ntreatments (n = 20 per species). Weeks show time since the start of the experiment, adjusted to\naccount for staggered start times. Survival probability is estimated from observed data\nusing a Kaplan Meier survival function. Letters in the legend (a, b, c, d) show species\nwhich are significantly different (post-hoc Tukey HSD test, p < 0.05).") +
  theme_pubclean() +
  custom_colors + 
  custom_colors_fill +
  theme(text = element_text(family = "serif", size = 10),
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 7, face = "italic"),
        strip.text.x = element_text(size = 10),
        plot.tag = element_text(family = "serif"),
        plot.caption = element_text(hjust = 0,
                                    family = "serif",
                                    #face = "bold",
                                    size = 10))
```

### Ambient Heatwave  {.hidden}

```{r}
# KM curve - Heatwave

#define colors
#names(myColorsPaired) <- levels(Data_All_hw$Legend)
names(myColorsDark) <- levels(Data_All_amb_hw$Legend)
names(myColorsLight) <- levels(Data_All_amb_hw$Legend)
custom_colors <- scale_colour_manual(values = myColorsLight)
custom_colors_fill <- scale_fill_manual(values = myColorsDark)

#Kaplan Meier Survival Curve - combined
km_species_fit <- survfit(Surv(Week, Dead_Count)~Legend, data = Data_All_amb_hw)

autoplot(km_species_fit) +
  # scale_fill_brewer(palette = "Paired") +
  # scale_color_brewer(palette = "Paired") +
  scale_x_continuous(limits = c(0, 36), breaks = seq(0 , 36, by = 2)) +
  #ylim(0,1) +
  annotate("rect",
           xmin = 7, xmax = 8,
           ymin = 0, ymax = 1,
           color = "red",
           fill = "red",
           linetype = "dashed",
           size = 0.8,
           alpha = 0.1) +
  geom_text(label = "heatwave",
            x = 5, y = 0.05, color = "red", size = 3.5, family = "serif") +
  xlab("Weeks") +
  ylab("Survival Probability") +
  # geom_text(label = "ponderosa pine",
  #           x = 13, y = 0.05, color = "black", size = 3, family = "serif") +
  # geom_text(label = "limber pine",
  #           x = 33, y = 0.05, color = "black", size = 3, family = "serif") +
  # geom_text(label = "Pinus ponderosa",
  #           x = 13, y = 0.05, color = "black", size = 3, family = "serif", fontface = "italic") +
  # geom_text(label = "Pinus flexilis",
  #           x = 33, y = 0.05, color = "black", size = 3, family = "serif", fontface = "italic") +
  labs(color = "", fill = "", tag = "(b)") +
  #labs(caption = "\nFIGURE S2a | Kaplan Meier Survival Curve of Juvenile Survival Probability under Drought and Heatwave. \nCurves shows the survival probability of each species under droughted and heatwave temperature \ntreatments (n = 20 per species). Weeks show time since the start of the experiment, adjusted to \naccount for staggered start times. Survival probability is estimated from observed data \nusing a Kaplan Meier survival function. Letters in the legend (a, b, c, d) show species \nwhich are significantly different (post-hoc Tukey HSD test, p < 0.05).") +
  theme_pubclean() +
  custom_colors +
  custom_colors_fill +
  theme(text = element_text(family = "serif", size = 10),
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 7, face = "italic"),
        strip.text.x = element_text(size = 10),
        plot.tag = element_text(family = "serif"),
        plot.caption = element_text(hjust = 0,
                                    family = "serif",
                                    #face = "bold",
                                    size = 10))

```

### Combined

```{r}
# KM curve - Ambient and Heatwave

#define colors
names(myColorsPaired) <- levels(Data_All_Phase1$Legend)
custom_colors <- scale_colour_manual(values = myColorsPaired)
custom_colors_fill <- scale_fill_manual(values = myColorsPaired)

#Kaplan Meier Survival Curve - separated by treatment
km_treatment_fit <- survfit(Surv(Week, Dead_Count)~Legend, data=Data_All_Phase1)
#summary(km_treatment_fit)


autoplot(km_treatment_fit) +
  # scale_fill_brewer(palette = "Paired") +
  # scale_color_brewer(palette = "Paired") +
  scale_x_continuous(breaks = seq(0 , 36, by = 2)) +
  annotate("rect",
           xmin = 7, xmax = 8,
           ymin = 0, ymax = 1,
           color = "red",
           fill = "red",
           linetype = "dashed",
           size = 0.8,
           alpha = 0.1) +
  geom_text(label = "heatwave",
            x = 5, y = 0.05, color = "red", size = 3.5, family = "serif") +
  xlab("Weeks") +
  ylab("Survival Probability") +
  # geom_text(label = "ponderosa pine",
  #           x = 13, y = 0.05, color = "black", size = 3, family = "serif") +
  # geom_text(label = "limber pine",
  #           x = 33, y = 0.05, color = "black", size = 3, family = "serif") +
  # geom_text(label = "Pinus ponderosa",
  #           x = 12.8, y = 0.05, color = "black", size = 3, family = "serif", fontface = "italic") +
  # geom_text(label = "Pinus flexilis",
  #           x = 33, y = 0.05, color = "black", size = 3, family = "serif", fontface = "italic") +
  labs(color = "", fill = "", tag = "(c)") +
  #labs(caption = "\nFIGURE S2b | Kaplan Meier Survival Curve of Juvenile Survival Probability under Drought. \nCurves shows the survival probability of each species under both ambient and heatwave temperature \ntreatments (n = 20 per species and treatment). Weeks show time since the start of the experiment, \nadjusted to account for staggered start times. Survival probability is estimated from observed data \nusing a Kaplan Meier survival function. Letters in the legend (a, b, c, d) show species \nwhich are significantly different (post-hoc Tukey HSD test, p < 0.05).") +
  theme_pubclean() +
  custom_colors +
  custom_colors_fill +
  theme(text = element_text(family = "serif", size = 10),
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 6, face = "italic"),
        strip.text.x = element_text(size = 10),
        plot.tag = element_text(family = "serif"),
        plot.caption = element_text(hjust = 0,
                                    family = "serif",
                                    #face = "bold",
                                    size = 10))
```

## Phase 2 - Hotter

### Hotter  {.hidden}

```{r}
# KM curve - Hotter

#define colors
#names(myColorsPaired) <- levels(Data_All_amb$Legend)
names(myColorsDark) <- levels(Data_All_hot$Legend)
names(myColorsLight) <- levels(Data_All_hot$Legend)
custom_colors <- scale_colour_manual(values = myColorsLight)
custom_colors_fill <- scale_fill_manual(values = myColorsDark)

#Kaplan Meier Survival Curve - combined
km_species_fit <- survfit(Surv(Week, Dead_Count)~Legend, data = Data_All_hot)

autoplot(km_species_fit) +
  scale_x_continuous(limits = c(0, 36), breaks = seq(0 , 36, by = 2)) +
  xlab("Weeks") +
  ylab("Survival Probability") +
  # geom_text(label = "ponderosa pine",
  #           x = 12, y = 0.05, color = "black", size = 3, family = "serif") +
  # geom_text(label = "limber pine",
  #           x = 33, y = 0.05, color = "black", size = 3, family = "serif") +
  # geom_text(label = "Pinus ponderosa",
  #           x = 12, y = 0.05, color = "black", size = 3, family = "serif", fontface = "italic") +
  # geom_text(label = "Pinus flexilis",
  #           x = 33, y = 0.05, color = "black", size = 3, family = "serif", fontface = "italic") +
  labs(color = "", fill = "", tag = "(a)") +
  #labs(caption = "\nFIGURE 1 | Kaplan Meier Survival Curve of Juvenile Survival Probability under Drought.\nCurves shows the survival probability of each species under droughted and ambient temperature\ntreatments (n = 20 per species). Weeks show time since the start of the experiment, adjusted to\naccount for staggered start times. Survival probability is estimated from observed data\nusing a Kaplan Meier survival function. Letters in the legend (a, b, c, d) show species\nwhich are significantly different (post-hoc Tukey HSD test, p < 0.05).") +
  theme_pubclean() +
  custom_colors + 
  custom_colors_fill +
  theme(text = element_text(family = "serif", size = 10),
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 7, face = "italic"),
        strip.text.x = element_text(size = 10),
        plot.tag = element_text(family = "serif"),
        plot.caption = element_text(hjust = 0,
                                    family = "serif",
                                    #face = "bold",
                                    size = 10))

```

### Hotter Heatwave  {.hidden}

```{r}
# KM curve - Heatwave

#define colors
#names(myColorsPaired) <- levels(Data_All_hw$Legend)
names(myColorsDark) <- levels(Data_All_hot_hw$Legend)
names(myColorsLight) <- levels(Data_All_hot_hw$Legend)
custom_colors <- scale_colour_manual(values = myColorsLight)
custom_colors_fill <- scale_fill_manual(values = myColorsDark)

#Kaplan Meier Survival Curve - combined
km_species_fit <- survfit(Surv(Week, Dead_Count)~Legend, data = Data_All_hot_hw)

autoplot(km_species_fit) +
  # scale_fill_brewer(palette = "Paired") +
  # scale_color_brewer(palette = "Paired") +
  scale_x_continuous(limits = c(0, 36), breaks = seq(0 , 36, by = 2)) +
  #ylim(0,1) +
  annotate("rect",
           xmin = 7, xmax = 8,
           ymin = 0, ymax = 1,
           color = "red",
           fill = "red",
           linetype = "dashed",
           size = 0.8,
           alpha = 0.1) +
  geom_text(label = "heatwave",
            x = 5, y = 0.05, color = "red", size = 3.5, family = "serif") +
  xlab("Weeks") +
  ylab("Survival Probability") +
  # geom_text(label = "ponderosa pine",
  #           x = 13, y = 0.05, color = "black", size = 3, family = "serif") +
  # geom_text(label = "limber pine",
  #           x = 33, y = 0.05, color = "black", size = 3, family = "serif") +
  # geom_text(label = "Pinus ponderosa",
  #           x = 13, y = 0.05, color = "black", size = 3, family = "serif", fontface = "italic") +
  # geom_text(label = "Pinus flexilis",
  #           x = 33, y = 0.05, color = "black", size = 3, family = "serif", fontface = "italic") +
  labs(color = "", fill = "", tag = "(b)") +
  #labs(caption = "\nFIGURE S2a | Kaplan Meier Survival Curve of Juvenile Survival Probability under Drought and Heatwave. \nCurves shows the survival probability of each species under droughted and heatwave temperature \ntreatments (n = 20 per species). Weeks show time since the start of the experiment, adjusted to \naccount for staggered start times. Survival probability is estimated from observed data \nusing a Kaplan Meier survival function. Letters in the legend (a, b, c, d) show species \nwhich are significantly different (post-hoc Tukey HSD test, p < 0.05).") +
  theme_pubclean() +
  custom_colors +
  custom_colors_fill +
  theme(text = element_text(family = "serif", size = 10),
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 7, face = "italic"),
        strip.text.x = element_text(size = 10),
        plot.tag = element_text(family = "serif"),
        plot.caption = element_text(hjust = 0,
                                    family = "serif",
                                    #face = "bold",
                                    size = 10))
```

### Combined

```{r}
# KM curve - Hotter and Heatwave

#define colors
names(myColorsPaired) <- levels(Data_All_Phase2$Legend)
custom_colors <- scale_colour_manual(values = myColorsPaired)
custom_colors_fill <- scale_fill_manual(values = myColorsPaired)

#Kaplan Meier Survival Curve - separated by treatment
km_treatment_fit <- survfit(Surv(Week, Dead_Count)~Legend, data=Data_All_Phase2)
#summary(km_treatment_fit)


autoplot(km_treatment_fit) +
  # scale_fill_brewer(palette = "Paired") +
  # scale_color_brewer(palette = "Paired") +
  scale_x_continuous(breaks = seq(0 , 36, by = 2)) +
  annotate("rect",
           xmin = 7, xmax = 8,
           ymin = 0, ymax = 1,
           color = "red",
           fill = "red",
           linetype = "dashed",
           size = 0.8,
           alpha = 0.1) +
  geom_text(label = "heatwave",
            x = 5, y = 0.05, color = "red", size = 3.5, family = "serif") +
  xlab("Weeks") +
  ylab("Survival Probability") +
  # geom_text(label = "ponderosa pine",
  #           x = 13, y = 0.05, color = "black", size = 3, family = "serif") +
  # geom_text(label = "limber pine",
  #           x = 33, y = 0.05, color = "black", size = 3, family = "serif") +
  # geom_text(label = "Pinus ponderosa",
  #           x = 12.8, y = 0.05, color = "black", size = 3, family = "serif", fontface = "italic") +
  # geom_text(label = "Pinus flexilis",
  #           x = 33, y = 0.05, color = "black", size = 3, family = "serif", fontface = "italic") +
  labs(color = "", fill = "", tag = "(c)") +
  #labs(caption = "\nFIGURE S2b | Kaplan Meier Survival Curve of Juvenile Survival Probability under Drought. \nCurves shows the survival probability of each species under both ambient and heatwave temperature \ntreatments (n = 20 per species and treatment). Weeks show time since the start of the experiment, \nadjusted to account for staggered start times. Survival probability is estimated from observed data \nusing a Kaplan Meier survival function. Letters in the legend (a, b, c, d) show species \nwhich are significantly different (post-hoc Tukey HSD test, p < 0.05).") +
  theme_pubclean() +
  custom_colors +
  custom_colors_fill +
  theme(text = element_text(family = "serif", size = 10),
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 6, face = "italic"),
        strip.text.x = element_text(size = 10),
        plot.tag = element_text(family = "serif"),
        plot.caption = element_text(hjust = 0,
                                    family = "serif",
                                    #face = "bold",
                                    size = 10))

```

# Death Date

```{r}
Data_Avg <- Data_All %>% 
  group_by(Phase, Species, Treatment_temp, Treatment_water, Week) %>%
  summarize(Dead_Count = sum(Dead_Count))

Data_Avg <- Data_Avg %>% 
  mutate(Live_Count = 20 - Dead_Count,
         Survivorship = ((Live_Count/20)*100))

Data_Avg %>% 
  filter(Treatment_water == "Drought", !is.na(Dead_Count)) %>% 
  group_by(Phase, Species, Treatment_temp) %>% 
  ggplot(aes(x = Week,
             y = Survivorship,
             color = Treatment_temp)) +
  geom_point() +
  geom_line() +
  facet_wrap(~Species) +
  theme_minimal()
```

```{r}
#boxplot
Data_All %>% 
  group_by(Phase, Species, Treatment_temp) %>% 
  arrange(Dead_Week) %>% 
  ggplot(aes(x = Dead_Week,
             y = Legend,
             fill = Treatment_temp)) +
  geom_boxplot() +
  facet_wrap(~Phase)+
  scale_x_continuous(limits = c(0, 40), breaks = seq(0 , 40, by = 2)) +
  xlab("Weeks to Mortality") +
  annotate("rect",
           xmin = 7, xmax = 8,
           ymin = 0, ymax = 10.25,
           color = "red",
           linetype = "dashed",
           size = 0.4,
           fill = "red",
           alpha = 0.3) +
  geom_text(label = "heatwave",
            x = 7.5, y = 10.5, color = "red", size = 3.5, family = "serif") +
  labs(fill = "") +
  theme_pubclean() +
  theme(text = element_text(family = "serif", size = 10),
        axis.text = element_text(size = 10),
        axis.text.y = element_text(face = "italic"),
        axis.title.y = element_blank(),
        legend.text = element_text(size = 10),
        strip.text.x = element_text(size = 10),
        plot.caption = element_text(hjust = 0,
                                    family = "serif",
                                    #face = "bold",
                                    size = 10))
```

# Porometer

```{r}
Data_All_p <- Data_All %>%
  mutate(Legend = ScientificName)
Data_All_p$Legend <- as.factor(Data_All_p$Legend)
Data_All_p <-
  transform(Data_All_p, Legend = factor(Legend, levels = c("Pinus ponderosa", "Pinus edulis", "Picea engelmannii", "Pseudotsuga menziesii", "Pinus flexilis")))
levels(Data_All_p$Legend)

label <- data.frame(x = c(6.0, 8.1, 12.5, 11.5, 16.4),
                    Legend = factor(c("Pinus ponderosa","Pinus edulis", "Picea engelmannii", "Pseudotsuga menziesii", "Pinus flexilis"), 
                                    levels = c("Pinus ponderosa", "Pinus edulis", "Picea engelmannii", "Pseudotsuga menziesii", "Pinus flexilis")))


#define custom color scale
myColorsPaired <- c("#6A3D9A", "#CAB2D6", "#FF7F00", "#FDBF6F",  "#33A02C", "#B2DF8A", "#E31A1C", "#FB9A99", "#1F78B4", "#A6CEE3")
myColorsDark <- c("#6A3D9A", "#FF7F00", "#33A02C", "#E31A1C", "#1F78B4")
myColorsLight <- c("#CAB2D6", "#FDBF6F", "#B2DF8A", "#FB9A99", "#A6CEE3")
names(myColorsPaired) <- levels(Data_All_p$Legend)
names(myColorsDark) <- levels(Data_All_p$Legend)
names(myColorsLight) <- levels(Data_All_p$Legend)

custom_colors <- scale_colour_manual(values = myColorsLight)
custom_colors_fill <- scale_fill_manual(values = myColorsDark)
```

## Phase 1 - Ambient

```{r}
#Graph
Data_All_p %>% 
  filter(Phase == 1,
         Treatment_water == "Drought") %>% 
  ggplot(aes(x = Week,
             y = Porometer,
             color = SpeciesID)) +
  geom_point() +
  #geom_line() +
  geom_segment(data = label,
               aes(x = x, xend = x,
                   y = 0, yend = 500),
               color = "black",
               linetype = "dashed",
               size = 0.8) +
  geom_segment(aes(x = 0, xend = 36,
                   y = 90, yend = 90),
               color = "grey",
               linetype = "solid",
               size = 0.8) +
  ylim(0, 500) +
  scale_x_continuous(breaks = seq(0 , 36, by = 4)) +
  facet_wrap(~Legend) +
  xlab("Weeks") +
  ylab(bquote("Stomatal Conductance  "~(mmol/m^2*s))) +
  #labs(caption = "\nFIGURE 4 | Stomatal Conductance of Juveniles under Droughted and Ambient Treatments.\nThe black dotted line shows the average time of permanent stomatal closure (n = 10 per species), when \nmeasurements were consistently at or below a minimum value of 90 mmol m-2s-1 (solid gray horizonal line).") +
  #custom_colors +
  theme_minimal() +
  theme(legend.position="none",
        text = element_text(family = "serif", size = 10),
        #axis.title.y = element_markdown(),
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 10),
        strip.text.x = element_text(size = 10, face = "italic"),
        plot.caption = element_text(hjust = 0,
                                    family = "serif",
                                    size = 10))
```

## Phase 2 - Hotter

```{r}
Data_All_p %>% 
  filter(Phase == 2,
         Treatment_water == "Drought") %>% 
  ggplot(aes(x = Week,
             y = Porometer,
             color = SpeciesID)) +
  geom_point() +
  #geom_line() +
  # geom_segment(data = label,
  #              aes(x = x, xend = x,
  #                  y = 0, yend = 500),
  #              color = "black",
  #              linetype = "dashed",
  #              size = 0.8) +
  geom_segment(aes(x = 0, xend = 36,
                   y = 90, yend = 90),
               color = "grey",
               linetype = "solid",
               size = 0.8) +
  ylim(0, 500) +
  scale_x_continuous(breaks = seq(0 , 36, by = 4)) +
  facet_wrap(~Legend) +
  xlab("Weeks") +
  ylab(bquote("Stomatal Conductance  "~(mmol/m^2*s))) +
  #labs(caption = "\nFIGURE 4 | Stomatal Conductance of Juveniles under Droughted and Ambient Treatments.\nThe black dotted line shows the average time of permanent stomatal closure (n = 10 per species), when \nmeasurements were consistently at or below a minimum value of 90 mmol m-2s-1 (solid gray horizonal line).") +
  #custom_colors +
  theme_minimal() +
  theme(legend.position="none",
        text = element_text(family = "serif", size = 10),
        #axis.title.y = element_markdown(),
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 10),
        strip.text.x = element_text(size = 10, face = "italic"),
        plot.caption = element_text(hjust = 0,
                                    family = "serif",
                                    size = 10))
```

# Weight

```{r}
Data_All_w <- Data_All %>%
  mutate(Legend = ScientificName)
Data_All_w$Legend <- as.factor(Data_All_w$Legend)
Data_All_w <-
  transform(Data_All_w, Legend = factor(Legend, levels = c("Pinus ponderosa", "Pinus edulis", "Picea engelmannii", "Pseudotsuga menziesii", "Pinus flexilis")))
levels(Data_All_w$Legend)

#filter data
Data_All_w <- Data_All_w %>% 
  filter(Treatment_water == "Drought",
         !is.na(WaterWeight_Calc))

#make label for inflection point segment
Data_All_w %>% 
  group_by(Legend) %>% 
  summarize(Stress_Week_Avg = mean(Stress_Week_Avg_Weight))

label <- data.frame(x = c(4.39, 4.32, 7.67, 7.82, 11.48),
                   Legend = factor(c("Pinus ponderosa","Pinus edulis", "Picea engelmannii", "Pseudotsuga menziesii", "Pinus flexilis"), 
                                   levels = c("Pinus ponderosa", "Pinus edulis", "Picea engelmannii", "Pseudotsuga menziesii", "Pinus flexilis")))

#define custom color scale
myColorsPaired <- c("#6A3D9A", "#CAB2D6", "#FF7F00", "#FDBF6F",  "#33A02C", "#B2DF8A", "#E31A1C", "#FB9A99", "#1F78B4", "#A6CEE3")
myColorsDark <- c("#6A3D9A", "#FF7F00", "#33A02C", "#E31A1C", "#1F78B4")
myColorsLight <- c("#CAB2D6", "#FDBF6F", "#B2DF8A", "#FB9A99", "#A6CEE3")
names(myColorsPaired) <- levels(Data_All_w$Legend)
names(myColorsDark) <- levels(Data_All_w$Legend)
names(myColorsLight) <- levels(Data_All_w$Legend)

custom_colors <- scale_colour_manual(values = myColorsLight)
custom_colors_fill <- scale_fill_manual(values = myColorsDark)
```

## Phase 1 - Ambient

```{r}
#Graph
Data_All_w %>% 
  filter(Phase == 1) %>% 
  ggplot(aes(x = Week,
             y = WaterWeight_Calc,
             color = SpeciesID)) +
  geom_point() +
  geom_line() +
  geom_segment(data = label,
               aes(x = x, xend = x,
                   y = 0, yend = 500),
               color = "black",
               linetype = "dashed",
               size = 0.8) +
  ylim(0, 500) +
  scale_x_continuous(breaks = seq(0 , 36, by = 4)) +
  facet_wrap(~Legend) +
  xlab("Weeks") +
  ylab("Water Weight (g)") +
  #labs(caption = "\nFIGURE 3 | Water Weight of Juveniles under Droughted and Ambient Treatments.\nCurves were calculated using individual weight data (n = 20 per species). Black dotted lines show the\naverage inflection point among all curves to show where concavity changes from concave up to concave down,\nreflecting the time when water availability transitioned from a non-limiting to a limiting resource.") +
  #custom_colors +
  theme_minimal() +
  theme(legend.position="none",
        text = element_text(family = "serif", size = 10),
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 10),
        strip.text.x = element_text(size = 10, face = "italic"),
        plot.caption = element_text(hjust = 0,
                                    family = "serif",
                                    #face = "bold",
                                    size = 10))
```

## Phase 2 - Hotter

```{r}
#Graph
Data_All_w %>% 
  filter(Phase == 2) %>% 
  ggplot(aes(x = Week,
             y = WaterWeight_Calc,
             color = SpeciesID)) +
  geom_point() +
  geom_line() +
  ylim(0, 500) +
  scale_x_continuous(breaks = seq(0 , 36, by = 4)) +
  facet_wrap(~Legend) +
  xlab("Weeks") +
  ylab("Water Weight (g)") +
  #labs(caption = "\nFIGURE 3 | Water Weight of Juveniles under Droughted and Ambient Treatments.\nCurves were calculated using individual weight data (n = 20 per species). Black dotted lines show the\naverage inflection point among all curves to show where concavity changes from concave up to concave down,\nreflecting the time when water availability transitioned from a non-limiting to a limiting resource.") +
  #custom_colors +
  theme_minimal() +
  theme(legend.position="none",
        text = element_text(family = "serif", size = 10),
        axis.text = element_text(size = 10),
        legend.text = element_text(size = 10),
        strip.text.x = element_text(size = 10, face = "italic"),
        plot.caption = element_text(hjust = 0,
                                    family = "serif",
                                    #face = "bold",
                                    size = 10))

```

# Percent Brown

```{r}
#filter NAs
Data_All_pb <- Data_All %>% 
  filter(!is.na(PercentBrown_Est)) %>% 
  mutate(PercentGreen_Est = 100 - PercentBrown_Est)

#average data 
Data_All_pb <- Data_All_pb %>%
  group_by(Phase, ScientificName, Species, Treatment_temp, Treatment_water, Week) %>%
  summarize(SampleSize_Weekly_PercentBrown = mean(SampleSize_Weekly_PercentBrown),
            Dead_Count = sum(Dead_Count),
            PercentBrown = round(mean(PercentBrown, na.rm = T), digits = 0),
            PercentBrown_Est = round(mean(PercentBrown_Est, na.rm = T), digits = 0),
            PercentGreen_Est = round(mean(PercentGreen_Est, na.rm = T), digits = 0),
            SD_PercentBrown = mean(SD_PercentBrown, na.rm = T))

#legend
Data_All_pb <- Data_All_pb %>%
  mutate(Legend = ScientificName)
Data_All_pb$Legend <- as.factor(Data_All_pb$Legend)
Data_All_pb <-
  transform(Data_All_pb, Legend = factor(Legend, levels = c("Pinus ponderosa", "Pinus edulis", "Picea engelmannii", "Pseudotsuga menziesii", "Pinus flexilis")))
levels(Data_All_pb$Legend)

#filter data
Data_All_pb <- Data_All_pb %>% 
  filter(Treatment_water == "Drought")

#define custom color scale
myColorsPaired <- c("#6A3D9A", "#CAB2D6", "#FF7F00", "#FDBF6F",  "#33A02C", "#B2DF8A", "#E31A1C", "#FB9A99", "#1F78B4", "#A6CEE3")
myColorsDark <- c("#6A3D9A", "#FF7F00", "#33A02C", "#E31A1C", "#1F78B4")
myColorsLight <- c("#CAB2D6", "#FDBF6F", "#B2DF8A", "#FB9A99", "#A6CEE3")
names(myColorsPaired) <- levels(Data_All_pb$Legend)
names(myColorsDark) <- levels(Data_All_pb$Legend)
names(myColorsLight) <- levels(Data_All_pb$Legend)

custom_colors <- scale_colour_manual(values = myColorsDark)
custom_colors_fill <- scale_fill_manual(values = myColorsDark)
```

## Phase 1 - Ambient

```{r}
#Graph
Data_All_pb %>% 
  filter(Phase == 1) %>% 
  ggplot(aes(x = Week,
             y = PercentGreen_Est,
             color = Legend)) +
  geom_point() +
  #scale_y_reverse() +
  geom_line() +
  geom_errorbar(aes(x = Week,
                    ymin = (PercentGreen_Est - SD_PercentBrown),
                    ymax = (PercentGreen_Est + SD_PercentBrown))) +
  # geom_segment(aes(x = 0, xend = 36,
  #                  y = 90, yend = 90),
  #              color = "grey",
  #              linetype = "solid",
  #              size = 0.8) +
  xlim(0, 36) +
  facet_grid(Legend~.) +
  xlab("Weeks") +
  ylab("Percent Green") +
  labs(caption = "FIGURE 5 | Ocular estimates of percent brown, averaged by each species per week") +
  custom_colors +
  theme_minimal() +
  theme(legend.position="none",
        text = element_text(family = "serif"),
        strip.text.y = element_text(angle = 0),
        plot.caption = element_text(hjust = 0,
                                    family = "serif",
                                    #face = "bold",
                                    size = 10))
```

## Phase 2 - Hotter

```{r}
#Graph
Data_All_pb %>% 
  filter(Phase == 2) %>% 
  ggplot(aes(x = Week,
             y = PercentGreen_Est,
             color = Legend)) +
  geom_point() +
  #scale_y_reverse() +
  geom_line() +
  geom_errorbar(aes(x = Week,
                    ymin = (PercentGreen_Est - SD_PercentBrown),
                    ymax = (PercentGreen_Est + SD_PercentBrown))) +
  # geom_segment(aes(x = 0, xend = 36,
  #                  y = 90, yend = 90),
  #              color = "grey",
  #              linetype = "solid",
  #              size = 0.8) +
  xlim(0, 36) +
  facet_grid(Legend~.) +
  xlab("Weeks") +
  ylab("Percent Green") +
  labs(caption = "FIGURE 5 | Ocular estimates of percent brown, averaged by each species per week") +
  custom_colors +
  theme_minimal() +
  theme(legend.position="none",
        text = element_text(family = "serif"),
        strip.text.y = element_text(angle = 0),
        plot.caption = element_text(hjust = 0,
                                    family = "serif",
                                    #face = "bold",
                                    size = 10))
```
